---
title: "Ancient woodland exploration"
author: "Ewan McHenry"
format:
  html:
    theme: cosmo
    reference-location: margin
    anchor-sections: true
    embed-resources: false
    toc: true
    code-fold: true
    default_chunk_options:
      message: false
    lightbox: true



knitr:
  opts_chunk:
    out.width: "100%"
    fig.height: 7  # Set your desired height
editor: visual
date: today
date-format: "D MMM, YYYY"

prefer-html: true

execute:
  echo: false
  warning: false
  cache: true
  cache_refresh: false
  message: false
  freeze: auto
---

```{r libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(sf) # for gis
library(terra)
library(units) # for set_units()
library(DT)
library(U.utilities) # devtools::install_github("EwanMcHenry/U.utilities")
library(RColorBrewer)
library(ggpubr)
library(plotly)
library(htmlwidgets)
library(htmltools)
library(RColorBrewer)
library(shiny)



```

```{r config, cache=F}
options(scipen = 999)  # Turn off scientific notation

#curation code
# source("Curation - NFI and NI woodland basemap.R")

constants <- list(hexdist.h = 5000,
                  hexdist.v = 5000,
                  focal_countries = c("England", "Scotland", "Wales", "N.Ireland"),
                  map.height = 5, 
                  map.width = 5,
                  scale.name = "Hex",
                  nucat_levels = c("ASNW", "SN ROY", "RAWS",
                                     "PAWS", "OTH ROY", 
                                     "Unknown AW",
                                     "SN LEPO", "OTH LEPO", 
                                     "Unknown LE"
                                    ),
                  high_levels = c("SN", "OTH", "Unknown"),
                  hexgrid_tol = 200

)

# colours

## pallete vis
# n_col <- 5  # colours per row
# 
# tibble(
#   index = seq_along(wt_palette),
#   colour = wt_palette
# ) %>%
#   mutate(
#     row = (index - 1) %/% n_col,
#     col = (index - 1) %% n_col
#   ) %>%
#   ggplot(aes(x = col, y = -row, fill = colour)) +
#   geom_tile(color = "white", linewidth = 1) +
#   geom_text(aes(label = index), size = 4) +
#   scale_fill_identity() +
#   coord_equal() +
#   theme_minimal() +
#   labs(title = "wt_palette") +
#   theme(
#     axis.title = element_blank(),
#     axis.text = element_blank(),
#     axis.ticks = element_blank()
#   )


# assign different colours to each AWI category, using the WT palette
nucat_colors <- setNames(wt_palette[c(1, 
                                      11, 10, 5,
                                      7,6,
                                      3,
                                      12, 14,
                                      9)], c("Total", constants$nucat_levels))
# high level colours
high_colors <- setNames(wt_palette[c(1, 11, 7, 9)], c("Total",constants$high_levels))






#plots config
colorCount = length(constants$nucat_levels)
# getPalette = colorRampPalette(brewer.pal(min(12, colorCount), "Paired"))
getPalette = colorRampPalette(wt_palette)


```

```{r load_data, echo=FALSE, message=FALSE, warning = FALSE}

load(paste0(gis.wd, "\\Data\\ancient woodland\\curated\\awi.RData"))
# load(paste0(gis.wd, "\\Data\\Designations\\SAC\\sac.RData"))
awi_df <- awi %>%
    as.data.frame()


countries0 = st_read("D:\\Users\\Ewan McHenry\\OneDrive - the Woodland Trust\\GIS\\Data\\administrative boundaries\\Countries\\5countries.v1.05.shp",
                     quiet = TRUE) %>% 
  st_transform(27700) %>% 
  st_buffer(0) %>% 
  st_simplify(dTolerance = 10) 
countries0$country.ha <-  st_area(countries0) %>% 
  set_units(value = "ha") %>%
  as.numeric()

countries <- countries0 %>% 
  rename(country = ctry17nm) %>%
  filter(country %in% constants$focal_countries ) %>%
  mutate(country = factor(country, levels = sort(unique(country))))

countries0 <- countries0 %>% 
    rename(country = ctry17nm) %>%
  st_simplify(dTolerance = 1000) # for plots - dont need high res

country.ha <- countries %>% 
  as.data.frame() %>% 
  select(country, country.ha) %>% 
  arrange(country) %>% 
  bind_rows(data.frame(country = "UK", country.ha = sum(.$country.ha)))   #add uk total
```

```{r awi_country_category_table, echo=FALSE, warning=FALSE, message=FALSE}
awi.country.category <- awi_df %>%
  group_by(Country, nu_cat) %>%
  summarise(awi.ha = sum(awi.ha, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = nu_cat, values_from = awi.ha, values_fill = 0) %>%
  filter(!is.na(Country))  # Ensures valid country entries

# Calculate UK totals and combine with country data
awi.country.category <- bind_rows(
  awi.country.category,
  awi.country.category %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE), Country = "UK")
)%>%
  select(Country, all_of(levels(awi$nu_cat)))

# Step 2: Reshape the data to have `awi` types as rows and `Country` as columns
awi_ha_table <- awi.country.category %>%
  pivot_longer(cols = -Country, names_to = "AWI_Type", values_to = "Area_ha") %>%
  pivot_wider(names_from = Country, values_from = Area_ha, values_fill = 0)  %>%
  replace_na(list(England = 0, Scotland = 0, Wales = 0, `N.Ireland` = 0, UK = 0))  # Ensure all NAs are replaced
# Calculate All AWI categories as a new row
total_awi_row <- awi_ha_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(AWI_Type = "All AWI categories") %>%
  select(AWI_Type, everything())
# Combine the new row with the original table
awi_ha_table <- bind_rows(awi_ha_table, total_awi_row) %>% 
  mutate(
    AWI_Type = factor(
      AWI_Type,
      levels = c(constants$nucat_levels, "All AWI categories")
    )
  )  %>% 
  arrange(AWI_Type)

# awi percentage cover by country
awi_per_table <- awi_ha_table 
for (i in 2:ncol(awi_per_table)){
  awi_per_table[,i] <- 100 * awi_ha_table[,i] / country.ha$country.ha[i-1] #so lazy!
}

awi_ha_long <- awi_ha_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "Area_ha") %>%
  filter(Area_ha > 0)  # Remove rows with 0 area

awi_per_long <- awi_per_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "percentage_cover")

awi_long <- awi_ha_long %>%
  left_join(awi_per_long, by = c("COUNTRY", "AWI_Type")) %>%
  mutate(percentage_cover = ifelse(is.na(percentage_cover), 0, percentage_cover))  # Replace NAs with 0

```

```{r make_hexgrid, cache=F}
hexgrid_file <- paste0("Data/hexgrid", constants$hexdist.h, "x", constants$hexdist.v, ".RData")

if (!file.exists(hexgrid_file)) {
  hexgrid <- st_make_grid(
    countries, 
    cellsize = c(constants$hexdist.h, constants$hexdist.v),
    square = FALSE
  ) %>% 
    st_sf() %>% 
    st_intersection(st_union(countries))

  hexgrid$hex.ha <- st_area(hexgrid) %>% set_units(value = "ha") %>% as.numeric()
  hexgrid$cell <- seq_len(nrow(hexgrid))

  save(hexgrid, file = hexgrid_file)
} else {
  load(hexgrid_file)
}

```

```{r st_aw_cover}
hexgrid_file <- paste0("Data/aw_hex_intersected", constants$hexdist.h, "x", constants$hexdist.v, ".rds")

if (!file.exists(hexgrid_file)) {
  aw_hex_intersected <- st_intersection(hexgrid, awi) %>%
    mutate(
      aw_ha_within_hex = st_area(.) %>%
        set_units("ha") %>%
        as.numeric()
    )
  
  # Save the object
  saveRDS(aw_hex_intersected, hexgrid_file)
} else {
  # Load the object if it already exists
  aw_hex_intersected <- readRDS(hexgrid_file)
}


hexgrid_aw <- aw_hex_intersected %>%  # Calculate the area of NFI polygons within each hex
  as.data.frame() %>%
  group_by(cell) %>% 
  summarise(
    total_aw_ha = sum(aw_ha_within_hex, na.rm = TRUE),  # Total woodland area within hex
    ASNW_ha = sum(aw_ha_within_hex[nu_cat == "ASNW"], na.rm = TRUE),  # ASNW area
    PAWS_ha = sum(aw_ha_within_hex[nu_cat == "PAWS"], na.rm = TRUE),  # PAWS area
    SN_LEPO_ha = sum(aw_ha_within_hex[nu_cat == "SN LEPO"], na.rm = TRUE),  # SN LEPO area
    OTH_LEPO_ha = sum(aw_ha_within_hex[nu_cat == "OTH LEPO"], na.rm = TRUE),  # OTH LEPO area
    Unknown_AW_ha = sum(aw_ha_within_hex[nu_cat == "Unknown AW"], na.rm = TRUE),  # Unknown AW area
    Unknown_LEPO_ha = sum(aw_ha_within_hex[nu_cat == "Unknown LE"], na.rm = TRUE),  # Unknown LEPO area
    RAWS_ha = sum(aw_ha_within_hex[nu_cat == "RAWS"], na.rm = TRUE),  # RAWS area
    SN_ROY_ha = sum(aw_ha_within_hex[nu_cat == "SN ROY"], na.rm = TRUE),  # SN ROY area
    OTH_ROY_ha = sum(aw_ha_within_hex[nu_cat == "OTH ROY"], na.rm = TRUE),  # OTH ROY area
        # Percentages relative to hex area
    total_aw_percentage = (total_aw_ha / hex.ha) * 100,
    ASNW_percentage = (ASNW_ha / hex.ha) * 100,
    PAWS_percentage = (PAWS_ha / hex.ha) * 100,
    SN_LEPO_percentage = (SN_LEPO_ha / hex.ha) * 100,
    OTH_LEPO_percentage = (OTH_LEPO_ha / hex.ha) * 100,
    Unknown_AW_percentage = (Unknown_AW_ha / hex.ha) * 100,
    Unknown_LEPO_percentage = (Unknown_LEPO_ha / hex.ha) * 100,
    RAWS_percentage = (RAWS_ha / hex.ha) * 100,
    SN_ROY_percentage = (SN_ROY_ha / hex.ha) * 100,
    OTH_ROY_percentage = (OTH_ROY_ha / hex.ha) * 100
  ) %>%
  ungroup() 

# Merge the calculated areas back onto the original hexgrid
hexgrid_final <- hexgrid %>%
  st_simplify(dTolerance = constants$hexgrid_tol) %>% 
  left_join(hexgrid_aw, by = "cell") %>%
  #replac all NA with 0
  mutate(across(where(is.numeric), ~replace_na(., 0)))
  

```

This is a living document. Any feedback welcomed, particularly if you want to suggest additions or report bugs etc (ewanmchenry\@woodlandtrust.org.uk).

# The AWI

The AWI used here is an amalgamation of the AWIs from each of the four countries. Notably, these use different methods of categorizing ancient woodland. I have attempted to unify these, based on my understanding of how the different categorizations. It gets a bit crazy[^1], hopefully this figure can help, or at least indicate the complexity of these datasets that some might initially think to be quite simple in their structure.

[^1]: I have a few more docs and notes on this, if you're interested, but here is not the place for that detail.

![How are the different AWI types categorised in the different countries' AWIs? How do I unify them to support comparison across coutnries.](images/AWI%20categories%20Sketch_2023-09-01%2015.23.31.excalidraw.sync-conflict-20240810-114956-OXL5OHY.svg)

# Area of Ancient Woodland Inventory (AWI) in Different Countries

::: panel-tabset
## Areas of AWI

```{r fig_awi_country_area, echo=FALSE}

# Create the ancient woodland area plot
aw_ha_plot <- awi_long %>%
  mutate(COUNTRY = factor(COUNTRY, levels = unique(COUNTRY)),
         AWI_Type = factor(AWI_Type, levels = rev(constants$nucat_levels))) %>%  # Reverse order for display
  ggplot(aes(fill = AWI_Type, y = Area_ha, x = COUNTRY,
             text = paste0("<b>", AWI_Type, " in ", COUNTRY, "</b><br>",
                           format(round(Area_ha, 0), big.mark = ","), " ha<br>",
                           round(percentage_cover, 1), "% of land"))
         ) +
  geom_bar(stat = "identity") +
  ggtitle("Ancient Woodland - Area") +
  theme_pubr() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Area (ha)") +
  scale_fill_manual(values = nucat_colors) +
  theme(legend.position = "right",
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(angle = 310, hjust = 0),
        axis.title.x = element_blank())

#save figure
ggsave("figures/awi_country_area.pdf", plot = aw_ha_plot, width = 178, dpi = 96 , 
       units = "mm")

# Convert to interactive plotly object
aw_ha_ggplotly <- ggplotly(aw_ha_plot, tooltip = "text", dynamicTicks = TRUE) %>%
   config(displayModeBar = FALSE)

aw_ha_ggplotly


```

```{r dt_awi_country_area_, echo=FALSE}

# Step 4: Create a datatable for display in the Quarto document
datatable(
  awi_ha_table %>%
  mutate(across(where(is.numeric), round, digits = 0)), 
  caption = "Area of different AWI categories (ha)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```

## AWI Percentage of Land Cover

```{r fig_awi_country_per, echo=FALSE}

aw_per_plot <- awi_long %>%
  mutate(COUNTRY = factor(COUNTRY, levels = unique(COUNTRY)),
         AWI_Type = factor(AWI_Type, levels = rev(constants$nucat_levels))) %>%  # Reverse order for display
  ggplot(aes(fill = AWI_Type, y = percentage_cover, x = COUNTRY,
             text = paste0("<b>", AWI_Type, " in ", COUNTRY, "</b><br>",
                           format(round(Area_ha, 0), big.mark = ","), " ha<br>",
                           round(percentage_cover, 1), "% of land"))
         ) +
  geom_bar(stat = "identity") +
  ggtitle("Ancient woodland land cover") +
  theme_pubr() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Land cover (%)") +
  scale_fill_manual(values = nucat_colors) +
  theme(legend.position = "right",
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(angle = 310, hjust = 0),
        axis.title.x = element_blank())

#save figure
ggsave("figures/awi_country_per.pdf", plot = aw_per_plot, width = 178, dpi = 96 , 
       units = "mm")

# Convert to interactive plotly object
aw_per_ggplotly <- ggplotly(aw_per_plot, tooltip = "text", dynamicTicks = TRUE) %>%
   config(displayModeBar = FALSE)

aw_per_ggplotly


```

```{r dt_awi_country_per, echo=FALSE}
datatable(
  awi_per_table %>%
  mutate(across(where(is.numeric), round, digits = 2)), 
  caption = "Cover of different AWI categories (% of land)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```
:::

# Mapping ancient woodland

In Scotland, there is more LEPO in the east, and more ASNW

There are two hotspots for PAWS

::: panel-tabset
## All AWI categories

```{r write_hexmap_shp}

final_awhexmap_output_path <- paste0("Outputs/", constants$hexdist.h, "x", constants$hexdist.v, "AWhexmap.gpkg")

# Only run if the file doesn't exist
if (!file.exists(final_awhexmap_output_path)) {
  # Make sure it's an sf object (skip st_as_sf if it's already sf)
  st_write(hexgrid_final, dsn = final_awhexmap_output_path, driver = "GPKG")
}


```

```{r AW_hexmap}

plot.col.lims <- c(0, quantile(hexgrid_final$total_aw_percentage[hexgrid_final$total_aw_percentage>0],
                               probs = 0.99))

awi_cover.hex.map <- map.ploter(to.plot = "total_aw_percentage",
                          fill.scale.title = expression(paste("AWI cover (%)",sep = "")),
                          main.title = "Ancient Woodland Inventory" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
) 

awi_cover.hex.map

# save figure
ggsave("figures/awi_cover_hexmap.pdf", plot = awi_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")

```

## ASNW

Ancient Semi-Natural Woodland (ASNW) is a subset of the AWI. It is defined as woodland that has been in existence since at least 1600 AD and has been influenced by human activities. ASNW is considered to be of high ecological value and is a priority habitat for conservation.

```{r hexmap_asnw}
plot.col.lims <- c(0, quantile(hexgrid_final$ASNW_percentage[hexgrid_final$ASNW_percentage>0],
                               probs = 0.995))

asnw_cover.hex.map <- map.ploter(to.plot = "ASNW_percentage",
                          fill.scale.title = expression(paste("ASNW cover (%)",sep = "")),
                          main.title = "Ancient Semi-Natural Woodland" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

asnw_cover.hex.map

# save figure
ggsave("figures/asnw_cover_hexmap.pdf", plot = asnw_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```

## PAWS

Plantation on Ancient Woodland Sites (PAWS) is a subset of the AWI. It is defined as an ancient woodland site, existing since 1600 AD, on which plantation woodland has been established. PAWS is a priority habitat for restoration.

```{r hexmap_paws}
plot.col.lims <- c(0, quantile(hexgrid_final$PAWS_percentage[hexgrid_final$PAWS_percentage>0],
                               probs = 0.995))

paws_cover.hex.map <- map.ploter(to.plot = "PAWS_percentage",
                          fill.scale.title = expression(paste("PAWS cover (%)",sep = "")),
                          main.title = "Plantations on Ancient Woodland Sites" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

paws_cover.hex.map

# save figure
ggsave("figures/paws_cover_hexmap.pdf", plot = paws_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```

## RAWS

Restored on Ancient Woodland Sites (RAWS) is a subset of the AWI. This woodland was formerly PAWS, but now has a canopy cover of at least 50% native species. It is important to note that RAWS is not a subset of ASNW, and that the process of restoration is ongoing on these sites: continued conservation management is needed for them to fulfill their ecological potential.

```{r hexmap_raws}
plot.col.lims <- c(0, quantile(hexgrid_final$RAWS_percentage[hexgrid_final$RAWS_percentage>0],
                               probs = 0.97))

raws_cover.hex.map <- map.ploter(to.plot = "RAWS_percentage",
                          fill.scale.title = expression(paste("RAWS cover (%)",sep = "")),
                          main.title = "Restored Ancient Woodland Sites" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

raws_cover.hex.map

# save figure
ggsave("figures/raws_cover_hexmap.pdf", plot = raws_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")

```

## Unknown Ancient Woodland

Ancient Woodland Inventory (AWI) data is not perfect, and the type of woodland on some AWI sites is uncertain. This map shows the proportion of hexagons covered by ancient woodland of unknown type.

```{r hexmap_unknown}
plot.col.lims <- c(0, quantile(hexgrid_final$Unknown_AW_percentage[hexgrid_final$Unknown_AW_percentage>0],
                               probs = 0.97))

unknown_cover.hex.map <- map.ploter(to.plot = "Unknown_AW_percentage",
                          fill.scale.title = expression(paste("Unknown cover (%)",sep = "")),
                          main.title = "Unknown Ancient Woodland" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

unknown_cover.hex.map

# save figure
ggsave("figures/unknown_cover_hexmap.pdf", plot = unknown_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```

## SN LEPO

Long Established Plantation Origin woodland (Semi-natural).

```{r hexmap_snlepo}
plot.col.lims <- c(0, quantile(hexgrid_final$SN_LEPO_percentage[hexgrid_final$SN_LEPO_percentage>0],
                               probs = 0.99))

snlepo_cover.hex.map <- map.ploter(to.plot = "SN_LEPO_percentage",
                          fill.scale.title = expression(paste("SN LEPO cover (%)",sep = "")),
                          main.title = "SN LEPO" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

snlepo_cover.hex.map

# save figure
ggsave("figures/snlepo_cover_hexmap.pdf", plot = snlepo_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```

## Other LEPO

Long Established Plantation Origin woodland (Non semi-natural).

```{r hexmap_otherlepo}
plot.col.lims <- c(0, quantile(hexgrid_final$OTH_LEPO_ha[hexgrid_final$OTH_LEPO_ha>0],
                               probs = 0.99))

otherlepo_cover.hex.map <- map.ploter(to.plot = "OTH_LEPO_ha",
                          fill.scale.title = expression(paste("Other LEPO cover (%)",sep = "")),
                          main.title = "Other LEPO" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

otherlepo_cover.hex.map

# save figure
ggsave("figures/otherlepo_cover_hexmap.pdf", plot = otherlepo_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")

```

## Unknown LEPO

AWI data is not perfect, the type of woodland on some LEPO sites is uncertain. This map shows the proportion of hexagons covered by LEPO of unknown type.

```{r hexmap_unknownlepo}
plot.col.lims <- c(0, quantile(hexgrid_final$Unknown_LEPO_percentage[hexgrid_final$Unknown_LEPO_percentage>0],
                               probs = 0.99))

unknownlepo_cover.hex.map <- map.ploter(to.plot = "Unknown_LEPO_percentage",
                          fill.scale.title = expression(paste("Unknown LEPO cover (%)",sep = "")),
                          main.title = "Unknown LEPO" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

unknownlepo_cover.hex.map

# save figure
ggsave("figures/unknownlepo_cover_hexmap.pdf", plot = unknownlepo_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```

## Roy SN

Roy is a sub-category of ancient woodland used in Scotland with evidence of existence in 1750, but with a potential of a breif unwooded interlude. Roy ancient woodlands appear on the 1750s Roy maps (providing historic evidence of woodland on sites extending beyond the maps on which other types of AWI are largely based), but may not appear as "woodland" on the c. 1850 1st edition OS. Many of these sites do however display many trees on the 1850s maps; the threshold of defining "woodland" on these is somewhat arbitrary. This map shows the proportion of hexagons covered by Roy ancient woodland with semi-natural characteristics.

```{r hexmap_roysn}
plot.col.lims <- c(0, quantile(hexgrid_final$SN_ROY_percentage[hexgrid_final$SN_ROY_percentage>0],
                               probs = 0.97))

roysn_cover.hex.map <- map.ploter(to.plot = "SN_ROY_percentage",
                          fill.scale.title = expression(paste("Roy SN cover (%)",sep = "")),
                          main.title = "Roy (Semi-Natural)" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

roysn_cover.hex.map

# save figure
ggsave("figures/roysn_cover_hexmap.pdf", plot = roysn_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")

```

## Roy Other

Roy is a sub-category of ancient woodland used in Scotland with evidence of existence in 1750, but with a potential of a brief unwooded interlude. Roy ancient woodlands appear on the 1750s Roy maps (providing historic evidence of woodland on sites extending beyond the maps on which other types of AWI are largely based), but may not appear as "woodland" on the c. 1850 1st edition OS. Many of these sites do however display many trees on the 1850s maps; the threshold of defining "woodland" on these is somewhat arbitrary. This map shows the proportion of hexagons covered by Roy ancient woodland with non semi-natural characteristics.

```{r hexmap_royother}
plot.col.lims <- c(0, quantile(hexgrid_final$OTH_ROY_percentage[hexgrid_final$OTH_ROY_percentage>0],
                               probs = 0.98))

royother_cover.hex.map <- map.ploter(to.plot = "OTH_ROY_percentage",
                          fill.scale.title = expression(paste("Roy Other cover (%)",sep = "")),
                          main.title = "Roy (Other)" ,
                          sub.title = NULL,
                          fillground = hexgrid_final,
                          background = countries0,
                          pltly.text = NULL,
                          transformation = "identity",
                          col.limits = plot.col.lims,
                          use.viridis = TRUE,
                          low.col = "white",
                          high.col = "red",
                          fill.line_size = 0.05,
                          fill.line_colour = NA,
                          background.fill = "grey90",
                          background.size = 0.05,
                          background.colour = "black",
                          n.breaks = 3,
                          round_to = 0.1,
                          just_pretty = T,
                          manual_breaks = NULL
)

royother_cover.hex.map

# save figure
ggsave("figures/royother_cover_hexmap.pdf", plot = royother_cover.hex.map, width = 178, dpi = 96 , 
       units = "mm")


```
:::

# Designated Ancient Woodland

## SSSI or SAC

```{r sac_sssi_awi_ha_table}
load(paste0(gis.wd, "\\Data\\Designations\\sac_sssi_awi.RData"))

sac_sssi_awi.country.category <- sac_sssi_awi %>%
    as.data.frame() %>%
  group_by(Country, nu_cat) %>%
  summarise(ha_sac_sssi_awi = sum(ha_sac_sssi_awi, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = nu_cat, values_from = ha_sac_sssi_awi, values_fill = 0) %>%
  filter(!is.na(Country))  # Ensures valid country entries
# Calculate UK totals and combine with country data
sac_sssi_awi.country.category <- bind_rows(
  sac_sssi_awi.country.category,
  sac_sssi_awi.country.category %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE), Country = "UK")
)%>%
  select(Country, all_of(levels(sac_sssi_awi$nu_cat)))
# Step 2: Reshape the data to have `awi` types as rows and `Country` as columns
sac_sssi_awi_ha_table <- sac_sssi_awi.country.category %>%
  pivot_longer(cols = -Country, names_to = "AWI_Type", values_to = "Area_ha") %>%
  pivot_wider(names_from = Country, values_from = Area_ha, values_fill = 0)  %>%
  replace_na(list(England = 0, Scotland = 0, Wales = 0, `N.Ireland` = 0, UK = 0))  # Ensure all NAs are replaced
# Calculate All AWI categories as a new row
total_sac_sssi_awi_row <- sac_sssi_awi_ha_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(AWI_Type = "All AWI categories") %>%
  select(AWI_Type, everything())
# Combine the new row with the original table
sac_sssi_awi_ha_table <- bind_rows(sac_sssi_awi_ha_table, total_sac_sssi_awi_row) %>% 
  arrange(factor(AWI_Type, levels = c(constants$nucat_levels, "All AWI categories")))

awi_per_sac_sssi_table <- sac_sssi_awi_ha_table 
for (i in 2:ncol(awi_per_sac_sssi_table)){
  awi_per_sac_sssi_table[,i] <- 100 * sac_sssi_awi_ha_table[,i] / awi_ha_table[,i] #so lazy!
}

sac_sssi_awi_ha_long <- sac_sssi_awi_ha_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "Area_ha") %>%
  filter(Area_ha > 0)  # Remove rows with 0 area

sac_sssi_awi_per_long <- awi_per_sac_sssi_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "percentage_sac_sssi")
sac_sssi_awi_long <- sac_sssi_awi_ha_long %>%
  left_join(sac_sssi_awi_per_long, by = c("COUNTRY", "AWI_Type")) %>%
  mutate(percentage_sac_sssi = ifelse(is.na(percentage_sac_sssi), 0, percentage_sac_sssi))  # Replace NAs with 0

```

### Area of SSSI or SAC Ancient Woodland

```{r fig_Area_of_SSI_or_SAC_ancient_woodland}
# Create the ancient woodland area plot
sac_sssi_aw_ha_plot <- sac_sssi_awi_long %>%
  mutate(COUNTRY = factor(COUNTRY, levels = unique(COUNTRY)),
         AWI_Type = factor(AWI_Type, levels = rev(constants$nucat_levels))) %>%  # Reverse order for display
  ggplot(aes(fill = AWI_Type, y = Area_ha, x = COUNTRY,
             text = paste0("<b>SSSI or SAC ", AWI_Type, " in ", COUNTRY, "</b><br>",
                           format(round(Area_ha, 0), big.mark = ","), " ha<br>",
                           round(percentage_sac_sssi, 1), "% of", AWI_Type, " in ", COUNTRY))
         ) +
  geom_bar(stat = "identity") +
  ggtitle("SSSI or SAC Ancient Woodland - Area") +
  theme_pubr() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Area (ha)") +
  scale_fill_manual(values = nucat_colors) +
  theme(legend.position = "right",
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(angle = 310, hjust = 0),
        axis.title.x = element_blank())
#save figure
ggsave("figures/sac_sssi_awi_country_area.pdf", plot = sac_sssi_aw_ha_plot, width = 178, dpi = 96 , 
       units = "mm")
# Convert to interactive plotly object
sac_sssi_aw_ha_ggplotly <- ggplotly(sac_sssi_aw_ha_plot, tooltip = "text", dynamicTicks = TRUE) %>%
   config(displayModeBar = FALSE)
sac_sssi_aw_ha_ggplotly
```

```{r dt_sac_sssi_awi_country_area_, echo=FALSE}
datatable(
  sac_sssi_awi_ha_table %>%
  mutate(across(where(is.numeric), round, digits = 0)), 
  caption = "Area of SSSI or SAC in different AWI categories (ha)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)

```

### Percentage of AWI that is SSSI or SAC

```{r dt_awi_per_sac_sssi, echo=FALSE}
datatable(
  awi_per_sac_sssi_table %>%
  mutate(across(where(is.numeric), round, digits = 2)), 
  caption = "SSSI or SAC cover of different AWI categories (%)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```

## SSSI

```{r sssi_aw_summarytable}
load(paste0(gis.wd, "\\Data\\Designations\\SSSI-ASSI\\sssi-awi.RData"))

sssi_awi.country.category <- sssi_awi %>%
    as.data.frame() %>%
  group_by(Country, nu_cat) %>%
  summarise(ha_sssi_awi = sum(ha_sssi_awi, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = nu_cat, values_from = ha_sssi_awi, values_fill = 0) %>%
  filter(!is.na(Country))  # Ensures valid country entries

# Calculate UK totals and combine with country data
sssi_awi.country.category <- bind_rows(
  sssi_awi.country.category,
  sssi_awi.country.category %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE), Country = "UK")
)%>%
  select(Country, all_of(levels(sssi_awi$nu_cat)))

# Step 2: Reshape the data to have `awi` types as rows and `Country` as columns
sssi_awi_ha_table <- sssi_awi.country.category %>%
  pivot_longer(cols = -Country, names_to = "AWI_Type", values_to = "Area_ha") %>%
  pivot_wider(names_from = Country, values_from = Area_ha, values_fill = 0)  %>%
  replace_na(list(England = 0, Scotland = 0, Wales = 0, `N.Ireland` = 0, UK = 0))  # Ensure all NAs are replaced
# Calculate All AWI categories as a new row
total_sssi_awi_row <- sssi_awi_ha_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(AWI_Type = "All AWI categories") %>%
  select(AWI_Type, everything())
# Combine the new row with the original table
sssi_awi_ha_table <- bind_rows(sssi_awi_ha_table, total_sssi_awi_row) %>% 
  arrange(factor(AWI_Type, levels = c(constants$nucat_levels, "All AWI categories")))

# awi percentage sssi by country
awi_per_sssi_table <- sssi_awi_ha_table 
for (i in 2:ncol(awi_per_sssi_table)){
  awi_per_sssi_table[,i] <- 100 * sssi_awi_ha_table[,i] / awi_ha_table[,i] #so lazy!
}

sssi_awi_ha_long <- sssi_awi_ha_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "Area_ha") %>%
  filter(Area_ha > 0)  # Remove rows with 0 area

sssi_awi_per_long <- awi_per_sssi_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "percentage_sssi")

sssi_awi_long <- sssi_awi_ha_long %>%
  left_join(sssi_awi_per_long, by = c("COUNTRY", "AWI_Type")) %>%
  mutate(percentage_sssi = ifelse(is.na(percentage_sssi), 0, percentage_sssi))  # Replace NAs with 0

```

*Note that, in N. Ireland, ASSI is the corresponding designation to SSSI.*

### Area of SSSI Ancient Woodland

```{r fig_Area_of_SSI_ancient_woodland}

# Create the ancient woodland area plot
sssi_aw_ha_plot <- sssi_awi_long %>%
  mutate(COUNTRY = factor(COUNTRY, levels = unique(COUNTRY)),
         AWI_Type = factor(AWI_Type, levels = rev(constants$nucat_levels))) %>%  # Reverse order for display
  ggplot(aes(fill = AWI_Type, y = Area_ha, x = COUNTRY,
             text = paste0("<b>SSSI ", AWI_Type, " in ", COUNTRY, "</b><br>",
                           format(round(Area_ha, 0), big.mark = ","), " ha<br>",
                           round(percentage_sssi, 1), "% of ", AWI_Type))
         ) +
  geom_bar(stat = "identity") +
  ggtitle("SSSI Ancient Woodland - Area") +
  theme_pubr() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Area (ha)") +
  scale_fill_manual(values = nucat_colors) +
  theme(legend.position = "right",
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(angle = 310, hjust = 0),
        axis.title.x = element_blank())

#save figure
ggsave("figures/sssi_awi_country_area.pdf", plot = sssi_aw_ha_plot, width = 178, dpi = 96 , 
       units = "mm")

# Convert to interactive plotly object
sssi_aw_ha_ggplotly <- ggplotly(sssi_aw_ha_plot, tooltip = "text", dynamicTicks = TRUE) %>%
   config(displayModeBar = FALSE)

sssi_aw_ha_ggplotly


```

```{r dt_sssi_awi_country_area_, echo=FALSE}
datatable(
  sssi_awi_ha_table %>%
  mutate(across(where(is.numeric), round, digits = 0)), 
  caption = "Area of SSSI in different AWI categories (ha)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)



```

### Percentage of AWI that is SSSI

```{r dt_awi_per_sssi, echo=FALSE}
datatable(
  awi_per_sssi_table %>%
  mutate(across(where(is.numeric), round, digits = 2)), 
  caption = "SSSI cover of different AWI categories (%)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```

## SAC

```{r sac_awi_ha_table}
load(paste0(gis.wd, "\\Data\\Designations\\SAC\\sac-awi.RData"))

sac_awi.country.category <- sac_awi %>%
    as.data.frame() %>%
  group_by(Country, nu_cat) %>%
  summarise(ha_sac_awi = sum(ha_sac_awi, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = nu_cat, values_from = ha_sac_awi, values_fill = 0) %>%
  filter(!is.na(Country))  # Ensures valid country entries
# Calculate UK totals and combine with country data
sac_awi.country.category <- bind_rows(
  sac_awi.country.category,
  sac_awi.country.category %>%
    summarise(across(where(is.numeric), sum, na.rm = TRUE), Country = "UK")
)%>%
  select(Country, all_of(levels(sac_awi$nu_cat)))


# Step 2: Reshape the data to have `awi` types as rows and `Country` as columns
sac_awi_ha_table <- sac_awi.country.category %>%
  pivot_longer(cols = -Country, names_to = "AWI_Type", values_to
 = "Area_ha") %>%
  pivot_wider(names_from = Country, values_from = Area_ha, values_fill = 0)  %>%
  replace_na(list(England = 0, Scotland = 0, Wales = 0, `N.Ireland` = 0, UK = 0))  # Ensure all NAs are replaced
# Calculate All AWI categories as a new row
total_sac_awi_row <- sac_awi_ha_table %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  mutate(AWI_Type = "All AWI categories") %>%
  select(AWI_Type, everything())
# Combine the new row with the original table
sac_awi_ha_table <- bind_rows(sac_awi_ha_table, total_sac_awi_row) %>% 
  arrange(factor(AWI_Type, levels = c(constants$nucat_levels, "All AWI categories")))

awi_per_sac_table <- sac_awi_ha_table 
for (i in 2:ncol(awi_per_sac_table)){
  awi_per_sac_table[,i] <- 100 * sac_awi_ha_table[,i] / awi_ha_table[,i] #so lazy!
}


sac_awi_ha_long <- sac_awi_ha_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "Area_ha") %>%
  filter(Area_ha > 0)  # Remove rows with 0 area

sac_awi_per_long <- awi_per_sac_table %>%
  filter(AWI_Type != "All AWI categories") %>%
  pivot_longer(cols = -AWI_Type, names_to = "COUNTRY", values_to = "percentage_sac")

sac_awi_long <- sac_awi_ha_long %>%
  left_join(sac_awi_per_long, by = c("COUNTRY", "AWI_Type")) %>%
  mutate(percentage_sac = ifelse(is.na(percentage_sac), 0, percentage_sac))  # Replace NAs with 0


```

### Area of SAC Ancient Woodland

```{r fig_Area_of_SAC_ancient_woodland}

# Create the ancient woodland area plot
sac_aw_ha_plot <- sac_awi_long %>%
  mutate(COUNTRY = factor(COUNTRY, levels = unique(COUNTRY)),
         AWI_Type = factor(AWI_Type, levels = rev(constants$nucat_levels))) %>%  # Reverse order for display
  ggplot(aes(fill = AWI_Type, y = Area_ha, x = COUNTRY,
             text = paste0("<b>SAC ", AWI_Type, " in ", COUNTRY, "</b><br>",
                           format(round(Area_ha, 0), big.mark = ","), " ha<br>",
                           round(percentage_sac, 1), "% of ", AWI_Type))
         ) +
  geom_bar(stat = "identity") +
  ggtitle("SAC Ancient Woodland - Area") +
  theme_pubr() +
  scale_x_discrete(name = NULL) +
  scale_y_continuous(name = "Area (ha)") +
  scale_fill_manual(values = nucat_colors) +
  theme(legend.position = "right",
        plot.title = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(angle = 310, hjust = 0),
        axis.title.x = element_blank())
#save figure
ggsave("figures/sac_awi_country_area.pdf", plot = sac_aw_ha_plot, width = 178, dpi = 96 , 
       units = "mm")

# Convert to interactive plotly object
sac_aw_ha_ggplotly <- ggplotly(sac_aw_ha_plot, tooltip = "text", dynamicTicks = TRUE) %>%
   config(displayModeBar = FALSE)

sac_aw_ha_ggplotly

```

```{r dt_sac_awi_country_area_, echo=FALSE}
datatable(
  sac_awi_ha_table %>%
  mutate(across(where(is.numeric), round, digits = 0)), 
  caption = "Area of SAC in different AWI categories (ha)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```

### Percentage of AWI that is SAC

```{r dt_awi_per_sac, echo=FALSE}
datatable(
  awi_per_sac_table %>%
  mutate(across(where(is.numeric), round, digits = 2)), 
  caption = "SAC cover of different AWI categories (%)",
  extensions = 'Buttons', 
  options = list(
    dom = 'Bfrtip',
    buttons = list('copy', 'excel'),
    paging = FALSE,  # Disable pagination
    scrollY = "400px",  # Enable scrolling
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = TRUE,
    searching = TRUE,  # Enable search/filter
    ordering = TRUE,  # Enable sorting
    pageLength = 10,  # Rows per page
    lengthMenu = c(10, 25, 50, 100)  # Rows per page options
  )
)
```

# Size of ancient woodland sites

An ask came from Nick Phillips about the size of AWI sites, primarily to inform policy around PAWS restoration by acting as evidence RE that commercial viability of e.g. harvest operations at these sites, suspecting that many were small, unviable for commercial forestry, and thus commercial forestery concerns may be less relevant for policy. Here I sought to investigate that speculation. An additional ask was made around the proportion on the public

In particular we were interested in the percentage of sites that were \< 3ha, which would essentially make them economically unviable for forestry operations.

Below are tables that summarise the site-area distributions of different AWI categories (high and detailed level), with info split by country. First, this is presented for sites (i.e. the number of sites distributed by their area) and then by (i.e. the combined area of sites of different areas). Obviously, while most sites may be small, more area can be within larger sites.

Notably, this is not related to habitat patch size, and should not be interpreted ecologically. Often AWI sites are contiguous with others, or other woodland that would effectively form part of the same habitat patch (while still being separate management units).

```{r f_cum_dens_awi}

# function to cumulative density plot of sites
f_cum_dens_awi_sites <- function(awi.data, category_col, size_col, colors) {
  
  # Make sure awi.data is a data frame
  df <- as.data.frame(awi.data)
  
  # Compute counts per category
  counts <- df %>%
    group_by(!!sym(category_col)) %>%
    summarise(n = n(), .groups = "drop")
  
  total_n <- sum(counts$n)
  
  # Build dynamic legend labels
  # "Total" first, then all other categories
  category_levels <- c("Total", as.character(counts[[category_col]]))
  
  # legendf labels
  legend_labels <- setNames(
    rep("", length(colors)),
    names(colors)
  )
  
  legend_labels["Total"] <- paste0("Total (N = ", scales::comma(total_n), ")")
  
  for(cat in setdiff(names(colors), "Total")) {
    n_cat <- counts$n[as.character(counts[[category_col]]) == cat]
    if(length(n_cat) == 0) n_cat <- 0
    legend_labels[cat] <- paste0(cat, " (N = ", scales::comma(n_cat), ")")
  }  
  # format legend
  n_legend_cols <- ifelse(length(colors) <= 4, 2, 3)
  
  
  # Plot ECDF
  ggplot(df, aes(x = !!sym(size_col), colour = !!sym(category_col))) +
    
    # Overall line
    stat_ecdf(
      aes(colour = "Total"),
      geom = "step",
      direction = "vh",
      linewidth = 1.2
    ) +
    
    # Category lines
    stat_ecdf(
      geom = "step",
      direction = "vh",
      linewidth = 1
    ) +
    
    scale_y_continuous(
      name = "Sites smaller (% of sites)",
      labels = scales::percent
    ) +
    scale_x_continuous(
      name = "Area of AWI site (ha)",
      expand = c(0, 0)
    ) +
    coord_cartesian(xlim = c(0, 12)) +
    scale_colour_manual(
      values = colors,
      breaks = names(colors),
      labels = legend_labels,
      name = "Category",
      guide = guide_legend(ncol = n_legend_cols, byrow = TRUE)
    ) +
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.key.width = unit(1.2, "lines")
          )
}

# function for cumulative density plot of area
f_cum_dens_awi_area <- function(awi.data, category_col, size_col, colors) {
  
  # Make sure awi.data is a data frame
  df <- as.data.frame(awi.data)
  
  # Compute total area per category
  area_sums <- df %>%
    group_by(!!sym(category_col)) %>%
    summarise(total_area = sum(!!sym(size_col), na.rm = TRUE), .groups = "drop")
  
  total_area_all <- sum(area_sums$total_area)
  
    # legendf labels
legend_labels <- setNames(
  rep("", length(colors)),
  names(colors)
)

legend_labels["Total"] <- paste0(
  "Total (Area = ", scales::comma(total_area_all), " ha)"
)

for(cat in setdiff(names(colors), "Total")) {
  
  area_cat <- area_sums$total_area[
    as.character(area_sums[[category_col]]) == cat
  ]
  
  if(length(area_cat) == 0) area_cat <- 0
  
  legend_labels[cat] <- paste0(
    cat, " (Area = ", scales::comma(area_cat), " ha)"
  )
}
  n_legend_cols <- ifelse(length(colors) <= 4, 2, 3)
  
  # Calculate cumulative area percentages
# categories
  df_cat <- df %>%
  group_by(!!sym(category_col)) %>%
  arrange(!!sym(size_col), .by_group = TRUE) %>%
  mutate(
    cum_area = cumsum(!!sym(size_col)),
    cum_area_perc = cum_area / sum(!!sym(size_col), na.rm = TRUE)
  ) %>%
  ungroup()
## total
df_total <- df %>%
  arrange(!!sym(size_col)) %>%
  mutate(
    cum_area = cumsum(!!sym(size_col)),
    cum_area_perc = cum_area / sum(!!sym(size_col), na.rm = TRUE),
    !!category_col := "Total"
  )
## combine
df_plot <- bind_rows(df_cat, df_total)

  # Plot cumulative area percentage
ggplot(df_plot, aes(x = !!sym(size_col), y = cum_area_perc, colour = !!sym(category_col))) +
  
  geom_step(direction = "vh", linewidth = 1) +
  
  scale_y_continuous(
    name = "Cumulative area \n(% of category's area)",
    labels = scales::percent
  ) +
  scale_x_continuous(
    name = "Area of AWI site (ha)",
    expand = c(0, 0)
  ) +
  coord_cartesian(xlim = c(0, 90)) +
  
  scale_colour_manual(
    values = colors,
    breaks = names(colors),
    labels = legend_labels,
    name = "Category",
    guide = guide_legend(ncol = n_legend_cols, byrow = TRUE)
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(1.2, "lines")
  )
}

```
```{r f_plot_dist_awi_sites, echo=FALSE}
plot_awi_distribution <- function(data,
                                  x_var,
                                  y_var = awi.ha,
                                  filter_var = NULL,
                                  filter_val = NULL,
                                  colours = nucat_colors) {
  
  # Optional filtering
  if (!is.null(filter_var) && !is.null(filter_val)) {
    data <- data %>%
      filter({{ filter_var }} %in% filter_val)
  }
  
  # Identify outliers
  outliers <- data %>%
    group_by({{ x_var }}) %>%
    mutate(
      q1 = quantile({{ y_var }}, 0.25, na.rm = TRUE),
      q3 = quantile({{ y_var }}, 0.75, na.rm = TRUE),
      iqr = q3 - q1,
      outlier = {{ y_var }} < (q1 - 1.5 * iqr) | {{ y_var }} > (q3 + 1.5 * iqr)
    ) %>%
    filter(outlier)
  
  # Plot
  ggplot(data, aes(x = {{ x_var }}, y = {{ y_var }}, fill = {{ x_var }})) +
    geom_jitter(
      data = outliers,
      aes(colour = {{ x_var }}),
      width = 0.3, height = 0,
      size = 1, alpha = 0.4
    ) +
    geom_boxplot(
      width = 0.1,
      outlier.shape = NA,
      fill = "white", colour = "black", size = 0.5
    ) +
    scale_fill_manual(values = colours) +
    scale_colour_manual(values = colours) +
    scale_x_discrete(name = NA) +
    scale_y_continuous(name = "Area of AWI site (ha)") +
    theme_minimal() +
    labs(title = "Distribution of AWI site areas by category",
         subtitle = "Boxplot of size area, with outliers as points") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 310, hjust = 0)
    )
}

```

::::::::::::::: panel-tabset
## High Level Categories

```{r dt_awi_site_size_high_level, echo=FALSE}

# tables

awi_cum_area_high_country <- awi_df %>%
  group_by(highlvl_SN_OTH, Country ) %>%
  arrange(awi.ha) %>%
  mutate(cum_area = cumsum(awi.ha),
         cum_area_perc = cum_area / sum(awi.ha, na.rm = TRUE)) %>% 
  summarise(total_area_ha = sum(awi.ha, na.rm = TRUE),
            n_sites = n(),
            median_size_ha = median(awi.ha, na.rm = TRUE),
            mean_size_ha = mean(awi.ha, na.rm = TRUE),
            site_area_at_25pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.25))],
            site_area_at_50pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.5))],
            site_area_at_75pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.75))],
            
            percentage_sites_under_3ha = 100 * sum(awi.ha < 3, na.rm = TRUE) / n(),
            percentage_sites_under_10ha = 100 * sum(awi.ha < 10, na.rm = TRUE) / n(),
            percentage_sites_under_20ha = 100 * sum(awi.ha < 20, na.rm = TRUE) / n(),
            percentage_sites_under_50ha = 100 * sum(awi.ha < 50, na.rm = TRUE) / n(),
            
            # percentage of area within sites 
            per_area_under_3ha = 100 * sum(awi.ha[awi.ha < 3], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_10ha = 100 * sum(awi.ha[awi.ha < 10], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_20ha = 100 * sum(awi.ha[awi.ha < 20], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_50ha = 100 * sum(awi.ha[awi.ha < 50], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE)

  )

awi_cum_area_high_uk <- awi_df %>%
    group_by(highlvl_SN_OTH ) %>%
    arrange(awi.ha) %>%
    mutate(cum_area = cumsum(awi.ha),
           cum_area_perc = cum_area / sum(awi.ha, na.rm = TRUE)) %>% 
  summarise(total_area_ha = sum(awi.ha, na.rm = TRUE),
            n_sites = n(),
            median_size_ha = median(awi.ha, na.rm = TRUE),
            mean_size_ha = mean(awi.ha, na.rm = TRUE),
            site_area_at_25pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.25))],
            site_area_at_50pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.5))],
            site_area_at_75pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.75))],
            
            percentage_sites_under_3ha = 100 * sum(awi.ha < 3, na.rm = TRUE) / n(),
            percentage_sites_under_10ha = 100 * sum(awi.ha < 10, na.rm = TRUE) / n(),
            percentage_sites_under_20ha = 100 * sum(awi.ha < 20, na.rm = TRUE) / n(),
            percentage_sites_under_50ha = 100 * sum(awi.ha < 50, na.rm = TRUE) / n(),
            
            # percentage of area within sites 
            per_area_under_3ha = 100 * sum(awi.ha[awi.ha < 3], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_10ha = 100 * sum(awi.ha[awi.ha < 10], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_20ha = 100 * sum(awi.ha[awi.ha < 20], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_50ha = 100 * sum(awi.ha[awi.ha < 50], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE)

  ) %>% 
  mutate(Country = "UK")

awi_cum_area_high_all <-  awi_cum_area_high_country %>% 
  rbind(awi_cum_area_high_uk)

awi_cum_area_high_all %>%
  select(highlvl_SN_OTH, Country, n_sites, mean_size_ha, median_size_ha, 
         site_area_at_25pct_cum_area, site_area_at_50pct_cum_area , site_area_at_75pct_cum_area, 
         percentage_sites_under_3ha, percentage_sites_under_10ha, percentage_sites_under_20ha, percentage_sites_under_50ha, 
         per_area_under_3ha, per_area_under_10ha, per_area_under_20ha, per_area_under_50ha,
         total_area_ha) %>%
  rename(Category = highlvl_SN_OTH,
         Country = Country,
         'Median area (ha)' = median_size_ha,
         'Mean area (ha)' = mean_size_ha,
         '% of sites < 3ha' = percentage_sites_under_3ha,
         '% of sites < 10ha' = percentage_sites_under_10ha,
         '% of sites < 20ha' = percentage_sites_under_20ha,
         '% of sites < 50ha' = percentage_sites_under_50ha,
         'Number of sites' = n_sites,
         'Total Area (ha)' = total_area_ha,
         '% of total area in sites < 3ha' = per_area_under_3ha,
         '% of total area in sites < 10ha' = per_area_under_10ha,
         '% of total area in sites < 20ha' = per_area_under_20ha,
         '% of total area in sites < 50ha' = per_area_under_50ha,
         'Site size at 25% cumulative area (ha)' = site_area_at_25pct_cum_area,
         'Site size at 50% cumulative area (ha)' = site_area_at_50pct_cum_area,
         'Site size at 75% cumulative area (ha)' = site_area_at_75pct_cum_area, 
         'Total area (ha)' = total_area_ha
         ) %>%
  # round numeric columns
  mutate(across(where(is.numeric), round, digits = 1)) %>%
  datatable(
    caption = "Size of AWI sites by high level category and country",
    extensions = 'Buttons', 
    options = list(
      dom = 'Bfrtip',
      buttons = list('copy', 'excel'),
      paging = FALSE,  # Disable pagination
      scrollY = "400px",  # Enable scrolling
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      searching = TRUE,  # Enable search/filter
      ordering = TRUE,  # Enable sorting
      pageLength = 10,  # Rows per page
      lengthMenu = c(20, 50, 100)  # Rows per page options
    )
  )

```

:::::::: panel-tabset
### All UK

::: panel-tabset

#### Site size distribution

```{r fig_awi_site_size_high_level_boxplot}
plot_awi_distribution(data = awi, x_var = highlvl_SN_OTH, y_var = awi.ha, colours = high_colors)
```

#### By number of sites

```{r fig_awi_site_size_high_level}
f_cum_dens_awi_sites(awi.data = awi, category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```

Unknown AWI types are typically much smaller, and are nearly always from Wales.

#### By cumulative area

```{r fig_awi_site_size_high_level_area}
f_cum_dens_awi_area(awi.data = awi, category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```
:::

### England

::: panel-tabset

#### Site size distribution

```{r fig_awi_site_size_high_level_boxplot_england}
plot_awi_distribution(data = awi %>% filter(Country == "England"), x_var = highlvl_SN_OTH, y_var = awi.ha, colours = high_colors)
```

#### By number of sites

```{r fig_awi_site_size_high_level_england}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "England"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```

In England, Semi-natural AWI sites are typically smaller than "Other" AWI sites (PAWS).

#### By cumulative area

```{r fig_awi_site_size_high_level_area_england}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "England"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```
:::

### N. Ireland

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_high_level_boxplot_ni}
plot_awi_distribution(data = awi %>% filter(Country == "N.Ireland"), x_var = highlvl_SN_OTH, y_var = awi.ha, colours = high_colors)
```
   
#### By number of sites

```{r fig_awi_site_size_high_level_ni}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "N.Ireland"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```

In N. Ireland unknown sites (of which there are few) are generally larger than the other categories.

#### By cumulative area

```{r fig_awi_site_size_high_level_area_ni}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "N.Ireland"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```
:::

### Scotland

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_high_level_boxplot_scotland}
plot_awi_distribution(data = awi %>% filter(Country == "Scotland"), x_var = highlvl_SN_OTH, y_var = awi.ha, colours = high_colors)
```

#### By number of sites

```{r fig_awi_site_size_high_level_scotland}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "Scotland"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```

#### By cumulative area

```{r fig_awi_site_size_high_level_area_scotland}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "Scotland"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```
:::

### Wales

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_high_level_boxplot_wales}
plot_awi_distribution(data = awi %>% filter(Country == "Wales"), x_var = highlvl_SN_OTH, y_var = awi.ha, colours = high_colors)
```

#### By number of sites

```{r fig_awi_site_size_high_level_wales}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "Wales"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```

#### By cumulative area

```{r fig_awi_site_size_high_level_area_wales}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "Wales"), category_col = "highlvl_SN_OTH", size_col = "awi.ha", colors = high_colors)
```
:::
::::::::

## Detailed Categories

```{r dt_awi_site_size_nucat, echo=FALSE}

awi_cum_area_nucat_country <- awi_df %>%
  group_by(nu_cat, Country ) %>%
  arrange(awi.ha) %>%
  mutate(cum_area = cumsum(awi.ha),
         cum_area_perc = cum_area / sum(awi.ha, na.rm = TRUE)) %>% 
  summarise(total_area_ha = sum(awi.ha, na.rm = TRUE),
            n_sites = n(),
            median_size_ha = median(awi.ha, na.rm = TRUE),
            mean_size_ha = mean(awi.ha, na.rm = TRUE),
            site_area_at_25pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.25))],
            site_area_at_50pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.5))],
            site_area_at_75pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.75))],
            
            percentage_sites_under_3ha = 100 * sum(awi.ha < 3, na.rm = TRUE) / n(),
            percentage_sites_under_10ha = 100 * sum(awi.ha < 10, na.rm = TRUE) / n(),
            percentage_sites_under_20ha = 100 * sum(awi.ha < 20, na.rm = TRUE) / n(),
            percentage_sites_under_50ha = 100 * sum(awi.ha < 50, na.rm = TRUE) / n(),
            # percentage of area within sites 
            per_area_under_3ha = 100 * sum(awi.ha[awi.ha < 3], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_10ha = 100 * sum(awi.ha[awi.ha < 10], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_20ha = 100 * sum(awi.ha[awi.ha < 20], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_50ha = 100 * sum(awi.ha[awi.ha < 50], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE)
            
  )

awi_cum_area_nucat_uk <- awi_df %>%
  group_by(nu_cat ) %>%
  arrange(awi.ha) %>%
  mutate(cum_area = cumsum(awi.ha),
         cum_area_perc = cum_area / sum(awi.ha, na.rm = TRUE)) %>% 
  summarise(total_area_ha = sum(awi.ha, na.rm = TRUE),
            n_sites = n(),
            median_size_ha = median(awi.ha, na.rm = TRUE),
            mean_size_ha = mean(awi.ha, na.rm = TRUE),
            site_area_at_25pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.25))],
            site_area_at_50pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.5))],
            site_area_at_75pct_cum_area = awi.ha[min(which(cum_area_perc >= 0.75))],

            percentage_sites_under_3ha = 100 * sum(awi.ha < 3, na.rm = TRUE) / n(),
            percentage_sites_under_10ha = 100 * sum(awi.ha < 10, na.rm = TRUE) / n(),
            percentage_sites_under_20ha = 100 * sum(awi.ha < 20, na.rm = TRUE) / n(),
            percentage_sites_under_50ha = 100 * sum(awi.ha < 50, na.rm = TRUE) / n(),
            
            # percentage of area within sites 
            per_area_under_3ha = 100 * sum(awi.ha[awi.ha < 3], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_10ha = 100 * sum(awi.ha[awi.ha < 10], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_20ha = 100 * sum(awi.ha[awi.ha < 20], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE),
            per_area_under_50ha = 100 * sum(awi.ha[awi.ha < 50], na.rm = TRUE) / sum(awi.ha, na.rm = TRUE)
            
  ) %>% 
  mutate(Country = "UK")

awi_cum_area_nucat_all <-  awi_cum_area_nucat_country %>%
  rbind(awi_cum_area_nucat_uk) %>% 
  arrange(factor(nu_cat, levels = c(constants$nucat_levels, "All AWI categories")))

awi_cum_area_nucat_all %>%
  select(nu_cat, Country, n_sites, mean_size_ha, median_size_ha, 
         site_area_at_25pct_cum_area, site_area_at_50pct_cum_area , site_area_at_75pct_cum_area, 
         percentage_sites_under_3ha, percentage_sites_under_10ha, percentage_sites_under_20ha, percentage_sites_under_50ha,
         per_area_under_3ha, per_area_under_10ha, per_area_under_20ha, per_area_under_50ha,
         total_area_ha) %>%
  rename(Category = nu_cat,
         Country = Country,
         'Median area (ha)' = median_size_ha,
         'Mean area (ha)' = mean_size_ha,
         '% of sites < 3ha' = percentage_sites_under_3ha,
         '% of sites < 10ha' = percentage_sites_under_10ha,
         '% of sites < 20ha' = percentage_sites_under_20ha,
         '% of sites < 50ha' = percentage_sites_under_50ha,
         'Number of sites' = n_sites,
         'Total Area (ha)' = total_area_ha,
         '% of total area in sites < 3ha' = per_area_under_3ha,
         '% of total area in sites < 10ha' = per_area_under_10ha,
         '% of total area in sites < 20ha' = per_area_under_20ha,
         '% of total area in sites < 50ha' = per_area_under_50ha,
         'Site size at 25% cumulative area (ha)' = site_area_at_25pct_cum_area,
         'Site size at 50% cumulative area (ha)' = site_area_at_50pct_cum_area,
         'Site size at 75% cumulative area (ha)' = site_area_at_75pct_cum_area, 
         'Total area (ha)' = total_area_ha
         ) %>%
  # round numeric columns
  mutate(across(where(is.numeric), round, digits = 1)) %>%
  datatable(
    caption = "Size of AWI sites by detailed category and country",
    extensions = 'Buttons',
    options = list(
      dom = 'Bfrtip',
      buttons = list('copy', 'excel'),
      paging = FALSE,  # Disable pagination
      scrollY = "400px",  # Enable scrolling
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      searching = TRUE,  # Enable search/filter
      ordering = TRUE,  # Enable sorting
      pageLength = 10,  # Rows per page
      lengthMenu = c(20, 50, 100)  # Rows per page options
    )
  )

```

:::::::: panel-tabset
### All UK

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_nucat_boxplot}
plot_awi_distribution(data = awi, x_var = nu_cat, y_var = awi.ha, colours = nucat_colors)
```

#### By number of sites

```{r fig_awi_site_size_nucat}

f_cum_dens_awi_sites(awi.data = awi, category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

#### By cumulative area

```{r fig_awi_site_size_nucat_area}
f_cum_dens_awi_area(awi.data = awi, category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

:::

### England

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_nucat_boxplot_england}
plot_awi_distribution(data = awi %>% filter(Country == "England"), x_var = nu_cat, y_var = awi.ha, colours = nucat_colors)
```

#### By number of sites

```{r fig_awi_site_size_nucat_england}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "England"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

#### By cumulative area

```{r fig_awi_site_size_nucat_area_england}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "England"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```
:::

### N. Ireland

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_nucat_boxplot_ni}
plot_awi_distribution(data = awi %>% filter(Country == "N.Ireland"), x_var = nu_cat, y_var = awi.ha, colours = nucat_colors)
```

#### By number of sites

```{r fig_awi_site_size_nucat_ni}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "N.Ireland"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

#### By cumulative area

```{r fig_awi_site_size_nucat_area_ni}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "N.Ireland"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```
:::

### Scotland

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_nucat_boxplot_scotland}
plot_awi_distribution(data = awi %>% filter(Country == "Scotland"), x_var = nu_cat, y_var = awi.ha, colours = nucat_colors)
```

#### By number of sites

```{r fig_awi_site_size_nucat_scotland}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "Scotland"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

#### By cumulative area

```{r fig_awi_site_size_nucat_area_scotland}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "Scotland"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```
:::

### Wales

::: panel-tabset
#### Site size distribution

```{r fig_awi_site_size_nucat_boxplot_wales}
plot_awi_distribution(data = awi %>% filter(Country == "Wales"), x_var = nu_cat, y_var = awi.ha, colours = nucat_colors)
```

#### By number of sites

```{r fig_awi_site_size_nucat_wales}
f_cum_dens_awi_sites(awi.data = awi %>% filter(Country == "Wales"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```

#### By cumulative area

```{r fig_awi_site_size_nucat_area_wales}
f_cum_dens_awi_area(awi.data = awi %>% filter(Country == "Wales"), category_col = "nu_cat", size_col = "awi.ha", colors = nucat_colors)
```
:::
::::::::
:::::::::::::::

## Where are the bigger AW sites?

```{r f_plot_point_map_size}
awi_centroids <- awi %>% 
  st_centroid() 

f_plot_awi_centroids_biggerthan <- function(size_threshold, category_col, colors) {

  plotdf <- awi_centroids %>% 
    filter(awi.ha > size_threshold)

  map.ploter(
    fillground = plotdf,
    to.plot = category_col,      # <- use the argument
    fill.scale.title = "AWI category",
    main.title = "Ancient Woodland Inventory",
    sub.title = paste("Sites >", size_threshold, "ha"),
    background = countries0,
    geom_type = "point",
    categorical = TRUE,
    colours = colors,            # <- use the argument
    background.fill = "grey90",
    background.size = 0.05,
    point.size = 0.6,
    point.alpha = 0.7,
    background.colour = "black"
  )
}


```


::: panel-tabset

## Sites > 3 ha
```{r}
f_plot_awi_centroids_biggerthan(size_threshold = 3, category_col = "nu_cat", colors = nucat_colors)
```

## Sites > 10 ha
```{r}
f_plot_awi_centroids_biggerthan(size_threshold = 10, category_col = "nu_cat", colors = nucat_colors)
```

## Sites > 50 ha
```{r}
f_plot_awi_centroids_biggerthan(size_threshold = 50, category_col = "nu_cat", colors = nucat_colors)
```

## Sites > 100 ha
```{r}
f_plot_awi_centroids_biggerthan(size_threshold = 100, category_col = "nu_cat", colors = nucat_colors)
```

:::


